Active Learning for SAT Solver Benchmarking {\large (extended and revised version)}

### 2025-02-12 paper_jar/paper.tex
Add clarification

As a baseline, Table~\ref{tab:regression} shows the prediction performance of two regression models, i.e., a random forest and a multilayer perceptron.
We employ these models to estimate a new solver's raw runtime as the prediction target.
To this end, we use 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes as inputs for the predictions.
We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers.

@@

Table~\ref{tab:timeout-prediction} shows the prediction performance for estimating whether a solver times out or not.
This binary prediction target represents a very strong discretization of runtimes.
We compare a variety of prediction models that pursue different learning paradigms.
Each is trained with either instance features only (second column) or instance and runtime features (third column).
As the evaluation metric for prediction performance, we use Matthews Correlation Coefficient (MCC)~\cite{gorodkin2004comparing, matthews1975comparison}, which reaches its maximum of~1 for a perfect prediction, is~0 for random guessing, and has a minimum of~-1.

Training the classification models with instance features only, a random forest performs best.
In contrast, a quadratic discriminant analysis performs poorly in this scenario but performs very well if runtime features are included.
However, the best prediction performance with all features is achieved by a stacking ensemble~\cite{wolpert1992stacked} out of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, \change{acts as a meta-model:
The prediction outputs of the ensemble members, i.e., quadratic discriminant analysis and random forests, become the prediction inputs of the meta-model, i.e., decision tree.
}

@@

Next, we use the runtimes labels from this discretization procedure as prediction targets.
Table~\ref{tab:cluster-prediction} shows the prediction performance for the best classification models from Table~\ref{tab:timeout-prediction} and $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
We train the prediction models either with 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).

@@

In this section, we evaluate our active learning framework.
First, we analyze and tune the different subroutines of our framework on the tuning dataset (cf.~Section~\ref{sec:eval:hyper}).
Next, we evaluate the best configurations on the full dataset (cf.~Section~\ref{sec:eval:full}).
Finally, we analyze the importance of instance families for our framework (cf.~Section~\ref{sec:eval:instance}).

>>>>

As a baseline, Table~\ref{tab:regression} shows the prediction performance of two regression models, i.e., a random forest and a multilayer perceptron.
We employ these models to estimate a new solver's raw runtime as the prediction target.
To this end, we use 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes as inputs for the predictions.
\change{For this experiment, the data is split into a training set (80\% of all data) on which the target runtime is known and a test set (20\% of all data) on which we predict the runtimes of the new solver.}
We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers \change{in Table~\ref{tab:regression}}.

@@

Table~\ref{tab:timeout-prediction} shows the prediction performance for estimating whether a solver times out or not.
This binary prediction target represents a very strong discretization of runtimes.
We compare a variety of prediction models that pursue different learning paradigms.
Each is trained with either instance features only (second column) or instance and runtime features (third column).
\change{For this experiment, the data is split into a training set (80\% of all data) on which the target label is known and a test set (20\% of all data) on which we predict the runtime label of the new solver.}
As the evaluation metric for prediction performance, we use Matthews Correlation Coefficient (MCC)~\cite{gorodkin2004comparing, matthews1975comparison}, which reaches its maximum of~1 for a perfect prediction, is~0 for random guessing, and has a minimum of~-1.

Training the classification models with instance features only \change{(as well as the target labels on the training set)}, a random forest performs best in predicting the target labels on the test set.
In contrast, a quadratic discriminant analysis performs poorly in this scenario but performs very well if runtime features are included.
However, the best prediction performance with all features \change{(as well as the target labels on the training set)} is achieved by a stacking ensemble~\cite{wolpert1992stacked} consisting of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, \change{acts as a meta-model:
The prediction outputs of the ensemble members, i.e., quadratic discriminant analysis and random forests, become the prediction inputs of the meta-model, i.e., decision tree.}

@@

Next, we use the runtimes labels from this discretization procedure as prediction targets.
Table~\ref{tab:cluster-prediction} shows the prediction performance for the best classification models from Table~\ref{tab:timeout-prediction} and $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
We train the prediction models either with 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
\change{For this experiment, the data is split into a training set (80\% of all data) on which the target cluster label is known and a test set (20\% of all data) on which we predict the cluster label of the new solver.}

@@

In this section, we evaluate our active learning framework.
\change{While we split the data into a fixed train and test set to evaluate performance in Section~\ref{sec:exdesign:disc-pred}, in this section, the target runtime labels of the new solver are only known for the experiments that have already been selected by active-learning.}
First, we analyze and tune the different subroutines of our framework on the tuning dataset (cf.~Section~\ref{sec:eval:hyper}).
Next, we evaluate the best configurations on the full dataset (cf.~Section~\ref{sec:eval:full}).
Finally, we analyze the importance of instance families for our framework (cf.~Section~\ref{sec:eval:instance}).
